### 2021-02-23 architecture.tex
sort > rank

The recommender consists of modules that perform specific subtasks,
each of which is described in detail in one of the following sections (see also \cref{fig:architecture}):
\begin{itemize}
\item \Cref{sec:graphifier}: Graphifier converts the input \gls{cnf} problem into a problem graph.
\item \Cref{sec:gcn}: \Gls{gcn} converts a problem graph into symbol embeddings.
\todo{MS: Kdyz embeddingy nebudou v preliminaries, je asi OK ctenare, ktery by o nich neslysel, zde trochu vydesit.
Ale v te prislusne podsekci tim zas zacit.}
\item \Cref{sec:output}: An output function \cite{Zhou2018} converts the symbol embeddings into a vector of symbol costs.
\todo{MS: ``output function \cite{Zhou2018}'' my nic nerika, mel bych to znat?\\
FB: Zhou2018 je review GNNs. Output function je termin, ktery se tam pouziva. Ta citace je nejspis tady nepotrebna.}
\item \Cref{sec:ranking}: Ranking converts the symbol costs into a symbol precedence.
It is only used in generating mode.
\todo{FB: Define ranking.}
\item \Cref{sec:training}: Loss function converts the symbol costs and a pair of precedences into a loss value.
It is only used in training mode.
\end{itemize}
\todo{MS: tohle zatim chapu jako predbezny nastrel. Leda by uz pred tim bylo vysvetleno:
embedding, \gls{gcn}/convolution layers, ... \\
FB: Staci predestrit, ze detaily vysvetlim pozdeji? MS: asi jo :)}

@@

\subsection{Ranking: From symbol costs to precedence}
\label{sec:ranking}

\todo[inline]{FB: Define ranking and argsort.}

@@

\begin{lemma}
Precedence cost $C$ is minimized by the precedence that ranks the symbols by their costs in non-increasing order:
% FB: The ranking terminology is from Fast Differentiable [...] Ranking.
$$
\argmin_\pi C(\pi) \in \argsort^- (c(1), \ldots, c(n))
$$
\end{lemma}

\begin{proof}
We prove the lemma by contradiction.
Let $\pi$ minimize $C$ and let $\pi$ not rank the costs in non-increasing order.
Then there exist $k < l$ such that $c(\pi(k)) < c(\pi(l))$.
Let $\bar{\pi}$ be a precedence obtained from $\pi$ by swapping the elements $k$ and $l$.
Then we obtain
%$\bar{\pi} = (\pi(1), \ldots, \pi(k-1), \pi(l), \pi(k+1), \ldots, \pi(l-1), \pi(k), \pi(l+1), \ldots, \pi(n))$.
\begin{align*}
\frac{C(\bar{\pi}) - C(\pi)}{Z_n}
&= kc(\bar{\pi}(k)) + lc(\bar{\pi}(l)) - kc(\pi(k)) - lc(\pi(l)) \\
&= kc(\pi(l)) + lc(\pi(k)) - kc(\pi(k)) - lc(\pi(l)) \\
&= k(c(\pi(l)) - c(\pi(k))) - l(c(\pi(l)) - c(\pi(k))) \\
&= (k-l) (c(\pi(l)) - c(\pi(k))) \\
&< 0
\end{align*}
\todo{FB: Isn't the inclusion of $Z_n$ too confusing?}
The final inequality is because $k-l < 0$ and $c(\pi(l)) - c(\pi(k)) > 0$.
Clearly, $Z_n > 0$ for any $n \geq 0$.
Thus, $C(\bar{\pi}) < C(\pi)$, which is contradicts the assumption that $\pi$ minimizes $C$. \qed
\end{proof}

Once the symbol costs are known,
producing a precedence that minimizes the precedence cost is cheap
because ranking is a fast operation.\todo{MS: can we not say ``sorting''? Also slightly earlier.\\
FB: This terminology is from the "Fast differentiable ranking" paper. Sorting rearranges the input vector. Ranking is argsort.}

@@

In \cref{sec:graphifier,sec:gcn,sec:output,sec:ranking} we described the structure of a recommender system that generates a precedence for an arbitrary input problem.
The efficacy of the recommender depends on the quality of the underlying symbol cost function.
In theory, the symbol cost function can assign the costs so that
ranking the symbols by their costs yields an optimum precedence.
Furthermore, all the information necessary to determine the optimum precedence is present in the graph representation of the input problem
thanks to the lossless property of the graph encoding mentioned before.
\todo{MS: Hlavne tady ale jakoby chybi druha pulky ty vety. Neco jako ``In practice, however, ...'' Nemyslis?} 
Our approach to defining an appropriate symbol cost function is based on statistical learning
from executions of an \gls{atp} on a set of problems with random precedences.

>>>>

The recommender consists of modules that perform specific subtasks,
each of which is described in detail in one of the following sections (see also \cref{fig:architecture}):
\begin{itemize}
\item \Cref{sec:graphifier}: Graphifier converts the input \gls{cnf} problem into a problem graph.
\item \Cref{sec:gcn}: \Gls{gcn} converts a problem graph into symbol embeddings.
\todo{MS: Kdyz embeddingy nebudou v preliminaries, je asi OK ctenare, ktery by o nich neslysel, zde trochu vydesit.
Ale v te prislusne podsekci tim zas zacit.}
\item \Cref{sec:output}: An output function \cite{Zhou2018} converts the symbol embeddings into a vector of symbol costs.
\todo{MS: ``output function \cite{Zhou2018}'' my nic nerika, mel bych to znat?\\
FB: Zhou2018 je review GNNs. Output function je termin, ktery se tam pouziva. Ta citace je nejspis tady nepotrebna.}
\item \Cref{sec:ranking}: Sorting converts the symbol costs into a symbol precedence.
It is only used in generating mode.
\item \Cref{sec:training}: Loss function converts the symbol costs and a pair of precedences into a loss value.
It is only used in training mode.
\end{itemize}
\todo{MS: tohle zatim chapu jako predbezny nastrel. Leda by uz pred tim bylo vysvetleno:
embedding, \gls{gcn}/convolution layers, ... \\
FB: Staci predestrit, ze detaily vysvetlim pozdeji? MS: asi jo :)}

@@

\subsection{Sorting: From symbol costs to precedence}
\label{sec:ranking}

@@

\begin{lemma}
Precedence cost $C$ is minimized by the precedence that sorts the symbols by their costs in non-increasing order:
$$
\argmin_\pi C(\pi) \in \argsort^- (c(1), \ldots, c(n))
$$
where $\argsort^-(x)$ is the set of all permutations $\rho$ that sort vector $x$ in non-increasing order ($x_{\rho(1)} \geq x_{\rho(2)} \geq \ldots \geq x_{\rho(n)}$).
\end{lemma}

\begin{proof}
We prove the lemma by contradiction.
Let $\pi$ minimize $C$ and let $\pi$ not sort the costs in non-increasing order.
Then there exist $k < l$ such that $c(\pi(k)) < c(\pi(l))$.
Let $\bar{\pi}$ be a precedence obtained from $\pi$ by swapping the elements $k$ and $l$.
Then we obtain
%$\bar{\pi} = (\pi(1), \ldots, \pi(k-1), \pi(l), \pi(k+1), \ldots, \pi(l-1), \pi(k), \pi(l+1), \ldots, \pi(n))$.
\begin{align*}
\frac{C(\bar{\pi}) - C(\pi)}{Z_n}
&= kc(\bar{\pi}(k)) + lc(\bar{\pi}(l)) - kc(\pi(k)) - lc(\pi(l)) \\
&= kc(\pi(l)) + lc(\pi(k)) - kc(\pi(k)) - lc(\pi(l)) \\
&= k(c(\pi(l)) - c(\pi(k))) - l(c(\pi(l)) - c(\pi(k))) \\
&= (k-l) (c(\pi(l)) - c(\pi(k))) \\
&< 0
\end{align*}
\todo{FB: Isn't the inclusion of $Z_n$ too confusing?}
The final inequality is because $k-l < 0$ and $c(\pi(l)) - c(\pi(k)) > 0$.
Clearly, $Z_n > 0$ for any $n \geq 0$.
Thus, $C(\bar{\pi}) < C(\pi)$, which is contradicts the assumption that $\pi$ minimizes $C$. \qed
\end{proof}

Once the symbol costs are known,
producing a precedence that minimizes the precedence cost is cheap
because sorting is a fast operation.

@@

In \cref{sec:graphifier,sec:gcn,sec:output,sec:ranking} we described the structure of a recommender system that generates a precedence for an arbitrary input problem.
The efficacy of the recommender depends on the quality of the underlying symbol cost function.
In theory, the symbol cost function can assign the costs so that
sorting the symbols by their costs yields an optimum precedence.
Furthermore, all the information necessary to determine the optimum precedence is present in the graph representation of the input problem
thanks to the lossless property of the graph encoding mentioned before.
\todo{MS: Hlavne tady ale jakoby chybi druha pulky ty vety. Neco jako ``In practice, however, ...'' Nemyslis?} 
Our approach to defining an appropriate symbol cost function is based on statistical learning
from executions of an \gls{atp} on a set of problems with random precedences.
